Northern Pikeminnow Predation on Juvenile Chinook Salmon in Deadwater Slough, Salmon River, Idaho and Potential Impacts to Local Populations

### 2021-12-29 analysis/paper/paper.Rmd
added very draft abstract

---
title: "Northern Pikeminnow Predation on Juvenile Chinook Salmon in Deadwater Slough, Salmon River, Idaho and Potential Impacts to Local Populations"
author:
  - Michael W. Ackerman:
      email: mike.ackerman@mthoodenvironmental.com
      institute: [mhe_mccall, biomark]
      correspondence: true
  - Nicholas J. Porter:
      email: Nicholas.Porter@merck.com
      institute: [biomark]
      correspondence: false
  - Kevin E. See:
      email: Kevin.See@dfw.wa.gov
      institute: [biomark, wdfw]
      correspondence: false
  - Michael S. Hall:
      email: Mike.Hall1@merck.com
      institute: [biomark]
      correspondence: false
  - Tara E. Blackman:
      email: tara.blackman@mthoodenvironmental.com
      institute: [mhe_sandy]
      correspondence: false
  - Richard A. Carmichael:
      email: Richard.Carmichael@merck.com
      institute: [biomark]
      correspondence: false
institute:
  - mhe_mccall: Mount Hood Environmental, PO Box 4282, McCall, Idaho, 83638, USA
  - biomark: Biomark, Inc., 705 South 8th Street, Boise, Idaho, 83702, USA
  - wdfw: Washington Department of Fish and Wildife, Fish Program, Science Division, 1111 Washington Street NE, Olympia, Washington, 98501, USA
  - mhe_sandy: Mount Hood Environmental, 39085 Pioneer Boulevard \#100 Mezzanine, Sandy, Oregon, 97055, USA 
date: "`r format(Sys.time(), '%d %B, %Y')`"
always_allow_html: true
output: 
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
      number_sections: FALSE
      toc: no
      always_allow_html: true
      fig_width: 6
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
    bookdown::html_document2:
      fig_caption: yes
      number_sections: FALSE
      toc: no
      fig_width: 6
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua      
fontsize: 12pt
mainfont: Times New Roman
bibliography: 
- references.bib      # Kevin's library
- AckermanLibrary.bib # Mike's library
csl: "../templates/american-fisheries-society.csl" # Insert path for the bib-style
abstract: |
  Abstract text here. 
keywords: |
  Northern Pikeminnow; Chinook Salmon; predation; mark-recapture; bioenergetics
highlights: |
  These are the highlights. 
---

@@

>>>>

---
title: "Northern Pikeminnow Predation on Juvenile Chinook Salmon in Deadwater Slough, Salmon River, Idaho and Potential Impacts to Local Populations"
author:
  - Michael W. Ackerman:
      email: mike.ackerman@mthoodenvironmental.com
      institute: [mhe_mccall, biomark]
      correspondence: true
  - Nicholas J. Porter:
      email: Nicholas.Porter@merck.com
      institute: [biomark]
      correspondence: false
  - Kevin E. See:
      email: Kevin.See@dfw.wa.gov
      institute: [biomark, wdfw]
      correspondence: false
  - Michael S. Hall:
      email: Mike.Hall1@merck.com
      institute: [biomark]
      correspondence: false
  - Tara E. Blackman:
      email: tara.blackman@mthoodenvironmental.com
      institute: [mhe_sandy]
      correspondence: false
  - Richard A. Carmichael:
      email: Richard.Carmichael@merck.com
      institute: [biomark]
      correspondence: false
institute:
  - mhe_mccall: Mount Hood Environmental, PO Box 4282, McCall, Idaho, 83638, USA
  - biomark: Biomark, Inc., 705 South 8th Street, Boise, Idaho, 83702, USA
  - wdfw: Washington Department of Fish and Wildife, Fish Program, Science Division, 1111 Washington Street NE, Olympia, Washington, 98501, USA
  - mhe_sandy: Mount Hood Environmental, 39085 Pioneer Boulevard \#100 Mezzanine, Sandy, Oregon, 97055, USA 
date: "`r format(Sys.time(), '%d %B, %Y')`"
always_allow_html: true
output: 
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
      number_sections: FALSE
      toc: no
      always_allow_html: true
      fig_width: 6
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
    bookdown::html_document2:
      fig_caption: yes
      number_sections: FALSE
      toc: no
      fig_width: 6
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua      
fontsize: 12pt
mainfont: Times New Roman
bibliography: 
- references.bib      # Kevin's library
- AckermanLibrary.bib # Mike's library
csl: "../templates/american-fisheries-society.csl" # Insert path for the bib-style
abstract: |
keywords: |
  Northern Pikeminnow; Chinook Salmon; predation; mark-recapture; bioenergetics
highlights: |
  These are the highlights. 
---

@@

## Abstract

Predation of emigrating juvenile salmonids by piscivorous fishes is perhaps an under-appreciated source of mortality hindering population recovery in the Columbia River basin. Reservoirs and tailraces associated with dams on the Snake and Columbia rivers have created favorable conditions for fishes known to consume juvenile salmonids, including the Northern Pikeminnow *Ptychocheilus oregonsis*. Indeed, Northern Pikeminnow have been estimated to consume 8% of the approximately 200 million juvenile salmonids emigration through those rivers, annually. Less is known about the interaction between upstream, river-dwelling Northern Pikeminnow and their impacts on salmonid recovery above the reservoir systems. In this study, we examine predation by Northern Pikeminnow on juvenile salmonids, focusing on Chinook Salmon *Oncorhynchus tshawytscha*, in a slow-water reach of the Salmon River, Idaho known as Deadwater Slough. The reach contains historically important overwinter rearing and spring emigration habitat for Chinook Salmon and other ESA-listed species, and was formed by the failure of a small mining dam in a reservoir in the late 19th century. We estimate the abundance of Northern Pikeminnow in Deadwater Slough during peak emigrations of Chinook salmon, and further, use a bioenergetics approach to estimate the consumption potential of the Northern Pikeminnow population to assess potential impacts to Chinook Salmon emigrants. Finally, we evaulate potential impacts to adult returns of local populations. ***We found that... (insert concluding findings here.*** It is likely that predation of juvenile Chinook Salmon in Deadwater Slough has a consequential impact on ESA-listed Chinook Salmon populations in the Upper Salmon River MPG hindering local recovery efforts.
